On the uniqueness and stability of dictionaries \\ for sparse representation of noisy signals

### 2019-05-09 manuscripts/IEEE/final_submission.tex
disc

There are still several challenges left open by our work. All conditions here guaranteeing uniqueness and stability of sparse representations have only been shown sufficient; it remains open, therefore, to extend them to necessary conditions, be they optimal lower bounds on required sample size, conditions on the support set hypergraphs of generating codes, or tolerable error bounds. On this last note, we remark that our constants have been derived for deterministic ``worst-case" noise, whereas the ``effective" noise might be significantly reduced when sampled from a concentrated distribution. It would be interesting to see how drastically all conditions can be tightened (to practical benefit) by requiring only probabilistic guarantees in this way, or in the spirit of the discussion on sample complexity following our proof of Thm.~\ref{DeterministicUniquenessTheorem}.

Another interesting question that remains is for which special cases is it efficient to check that a solution to Probs.~\ref{InverseProblem} or \ref{OptimizationProblem} is unique (and stable). Given the NP-hardness of computing sufficient conditions for checking this in general, are the corresponding problems also computationally hard?  Finally, since substituting the norm $\|\cdot\|_0$ with $\|\cdot\|_1$ yields the same optimal sparse codes for dictionaries with large enough lower bounds \cite{eldar2012compressed}, is there a version of Thm.~\ref{SLCopt} certifying that Prob.~\ref{OptimizationProblem} can be solved while inferring sparse codes in this way?  % [*** Incorporate this open problem with more words ***]

@@

%We close with some challenges left open by our work. First, there is the basic problem of proving optimal lower bounds on sufficient sample size $N$ guaranteeing the uniqueness and stability of sparse representations. The pigeonholing argument in the proof of Thm.~\ref{DeterministicUniquenessTheorem} is loose in that it assigns to each support $S \in \mathcal{H}$ a corresponding support $\overline S$ without consideration of previously matched pairs; although we doubt tightening in this way will lead to sub-exponential sample sizes. Sample complexity will also benefit from lowering the number of required supports, i.e. the size of regular uniform hypergraphs satisfying the SIP. The regularity requirement on support hypergraphs can be loosened to allow for certain non-regular hypergraphs satisfying constraints on the degree sequence of nodes; though we impose regularity for clarity of exposition. It is then natural to ask: what is the smallest hypergraph satisfying the required constraints forgiven $m$ and $k$? Finally, the error bounds can be improved by tightening the constant $C_1$, e.g. via an improvement in Lemma ?.

>>>>

@@

We close with some challenges left open by our work. All conditions here guaranteeing uniqueness and stability of sparse representations have only been shown sufficient; it remains open, therefore, to extend them to necessary conditions, be they optimal bounds on required sample size, conditions on the support set hypergraphs of generating codes, or tolerable error bounds. On this last note, we remark that our constants have been derived for deterministic ``worst-case" noise, whereas the ``effective" noise might be significantly reduced when sampled from a concentrated distribution. It would be interesting to see how drastically all conditions can be relaxed (to practical benefit) by requiring only probabilistic guarantees in this way, or in the spirit of the discussion on sample complexity following our proof of Thm.~\ref{DeterministicUniquenessTheorem}.

Another interesting question is for which special cases is it efficient to check that a solution to Prob.~\ref{InverseProblem} or \ref{OptimizationProblem} is unique (and stable). Considering that our sufficient conditions for checking this in general are NP-hard to compute, are the corresponding problems also hard?  Finally, since substituting the norm $\|\cdot\|_0$ with $\|\cdot\|_1$ yields the same optimal sparse codes for dictionaries with large enough lower bounds \cite{eldar2012compressed}, is there a version of Thm.~\ref{SLCopt} certifying that Prob.~\ref{OptimizationProblem} can be solved while inferring sparse codes in this way?  % [*** Incorporate this open problem with more words ***]
It is our hope that these remaining challenges peak the interest of the community enough to pick up where we have left off.
%We close with some challenges left open by our work. First, there is the basic problem of proving optimal lower bounds on sufficient sample size $N$ guaranteeing the uniqueness and stability of sparse representations. The pigeonholing argument in the proof of Thm.~\ref{DeterministicUniquenessTheorem} is loose in that it assigns to each support $S \in \mathcal{H}$ a corresponding support $\overline S$ without consideration of previously matched pairs; although we doubt tightening in this way will lead to sub-exponential sample sizes. Sample complexity will also benefit from lowering the number of required supports, i.e. the size of regular uniform hypergraphs satisfying the SIP. The regularity requirement on support hypergraphs can be loosened to allow for certain non-regular hypergraphs satisfying constraints on the degree sequence of nodes; though we impose regularity for clarity of exposition. It is then natural to ask: what is the smallest hypergraph satisfying the required constraints forgiven $m$ and $k$? Finally, the error bounds can be improved by tightening the constant $C_1$, e.g. via an improvement in Lemma ?.
